Fast and scalable Gaussian process modeling with applications to astronomical time series \vspace{-3\baselineskip}

### 2017-03-20 paper/ms.tex
Added TESS, SOAP references

To demonstrate the applicability of this model, we apply it to infer the
asteroseismic parameters of the giant star KIC~11615890 observed by the
\kepler\ Mission.
The goal of this example is to show that, even for a low signal-to-noise
dataset with a short baseline, it is possible to infer asteroseismic
parameters with formal uncertainties that are consistent with the parameters
inferred with a much larger dataset.
Looking forward to \tess\ \citep{Campante:2016}, we will infer
$\nu_\mathrm{max}$ and $\Delta\nu$ using only one month of \kepler\ data and
compare our results to the results inferred from the full 4 year baseline of
the \kepler\ mission.
For KIC~11615890, the published asteroseismic parameters based on several
years of \kepler\ observations are \citep{Pinsonneault:2014}
\begin{eqnarray}
    \nu_\mathrm{max} = 171.94 \pm 3.62 \,\mu\mathrm{Hz} \quad\mathrm{and}\quad
    \Delta\nu = 13.28 \pm 0.29 \,\mu\mathrm{Hz} \quad.
\end{eqnarray}
We randomly select a month-long segment of \kepler\ data, initialize our
\celerite\ model using a grid search in the parameter space, and then use
\emcee\ \citep{Foreman-Mackey:2013} to sample the joint posterior density for
the full set of parameters.
\Figure{astero-corner} shows the marginalized density for $\nu_\mathrm{max}$
and $\Delta\nu$ compared to the results from the literature.
\dfmtodo{Add some more discussion here.}

@@

Another source of variability that can be measured from time series
measurements of stars is rotation.
The inhomogeneous surface of the star (spots, plage, \etc) imprints itself as
quasiperiodic variations in photometric or spectroscopic observations
(\dfmtodo{CITE}).
It has been demonstrated that for light curves with nearly uniform sampling,
the empirical autocorrelation function provides a reliable estimate of the
rotation period of a star (\dfmtodo{CITE}) and that a GP model with a
quasiperiodic covariance function can be used to make probabilistic
measurements even with sparsely sampled data (R.~Angus, \etal\ in prep.).
The covariance function used for this type of analysis has the form
\begin{eqnarray}\eqlabel{sine2}
k(\tau) = A\,\exp\left(-\frac{\tau^2}{2\,\ell^2} -
    \Gamma\,\sin^2\left(\frac{\pi\,\tau}{P} \right) \right)
\end{eqnarray}
where $P$ is the period of the oscillation.
The key difference between this function and other quasiperiodic kernels is
that it is everywhere positive.
We can construct a simple \celerite\ covariance function with similar properties
as follows
\begin{eqnarray}\eqlabel{rot-kernel}
k(\tau) = \frac{a}{2+b}\,e^{-c\,\tau}\,\left[
    \cos\left(\frac{2\,\pi\,\tau}{P}\right) + (1 + b)
\right]
\end{eqnarray}
for $a>0$, $b>0$, and $c>0$.
The covariance function in \eq{rot-kernel} cannot exactly reproduce \eq{sine2}
but, since \eq{sine2} is only an effective model, \eq{rot-kernel} can be used
as a drop-in replacement for a substantial gain in computational efficiency.

>>>>

To demonstrate the applicability of this model, we apply it to infer the
asteroseismic parameters of the giant star KIC~11615890 observed by the
\kepler\ Mission.
The goal of this example is to show that, even for a low signal-to-noise
dataset with a short baseline, it is possible to infer asteroseismic
parameters with formal uncertainties that are consistent with the parameters
inferred with a much larger dataset.
Looking forward to \tess\ \citep{Ricker:2014,Campante:2016}, we will infer
$\nu_\mathrm{max}$ and $\Delta\nu$ using only one month of \kepler\ data and
compare our results to the results inferred from the full 4 year baseline of
the \kepler\ mission.
For KIC~11615890, the published asteroseismic parameters based on several
years of \kepler\ observations are \citep{Pinsonneault:2014}
\begin{eqnarray}
    \nu_\mathrm{max} = 171.94 \pm 3.62 \,\mu\mathrm{Hz} \quad\mathrm{and}\quad
    \Delta\nu = 13.28 \pm 0.29 \,\mu\mathrm{Hz} \quad.
\end{eqnarray}
We randomly select a month-long segment of \kepler\ data, initialize our
\celerite\ model using a grid search in the parameter space, and then use
\emcee\ \citep{Foreman-Mackey:2013} to sample the joint posterior density for
the full set of parameters.
\Figure{astero-corner} shows the marginalized density for $\nu_\mathrm{max}$
and $\Delta\nu$ compared to the results from the literature.
\dfmtodo{Add some more discussion here.}

@@

Another source of variability that can be measured from time series
measurements of stars is rotation.
The inhomogeneous surface of the star (spots, plage, \etc) imprints itself as
quasiperiodic variations in photometric or spectroscopic observations
\citep{Dumusque:2014}.
It has been demonstrated that for light curves with nearly uniform sampling,
the empirical autocorrelation function provides a reliable estimate of the
rotation period of a star (\dfmtodo{CITE}) and that a GP model with a
quasiperiodic covariance function can be used to make probabilistic
measurements even with sparsely sampled data (R.~Angus, \etal\ in prep.).
The covariance function used for this type of analysis has the form
\begin{eqnarray}\eqlabel{sine2}
k(\tau) = A\,\exp\left(-\frac{\tau^2}{2\,\ell^2} -
    \Gamma\,\sin^2\left(\frac{\pi\,\tau}{P} \right) \right)
\end{eqnarray}
where $P$ is the period of the oscillation.
The key difference between this function and other quasiperiodic kernels is
that it is everywhere positive.
We can construct a simple \celerite\ covariance function with similar properties
as follows
\begin{eqnarray}\eqlabel{rot-kernel}
k(\tau) = \frac{a}{2+b}\,e^{-c\,\tau}\,\left[
    \cos\left(\frac{2\,\pi\,\tau}{P}\right) + (1 + b)
\right]
\end{eqnarray}
for $a>0$, $b>0$, and $c>0$.
The covariance function in \eq{rot-kernel} cannot exactly reproduce \eq{sine2}
but, since \eq{sine2} is only an effective model, \eq{rot-kernel} can be used
as a drop-in replacement for a substantial gain in computational efficiency.
